How to use replication assignments for teaching integrity in empirical archaeology

### 2019-08-16 analysis/paper/paper.Rmd
improve the readme to make it more clear what this is, and what to do with it
Add Hope's commas 👓

abstract: |
  The value of new archaeological knowledge is strongly determined by how credible it is, and a key measure of scientific credibility is how replicable new results are. However, few archaeologists learn the skills necessary to conduct replication as part of their training. This means there is a gap between the ideals of archaeological science, and the skills we teach future researchers. Here we argue for replications as a core type of class assignment in archaeology courses to close this gap and establish a culture of replication and reproducibility. We review replication assignments in other fields, and describe how to implement a replication assignment suitable for many types of archaeology programs. We describe our experience with replication in an upper level undergraduate class on stone artefact analysis. Replication assignments can help archaeology programs give students skills that enable transparent and reproducible research.

@@

In his influential study of replication, sociologist of science Harry Collins argued that replication is at the core of scientific practice, writing “Replicability [...] is the Supreme Court of the scientific system” [@collins1992changing p. 19]. Like other observers of science, Collins claimed that in replication, private observations become communal facts, offering vital protection from error and fraud. In this paper we propose a new type of assignment for the archaeology classroom, the replication report, to better align the practice of teaching archaeology with the transparency and openness ideals of science [@nosek2015promoting]. The replication report assignment involves three four steps: 1) students analysing a published report to determine the main claims made by the authors of that report, 2) students obtaining the data used by the authors, 3) analysing that data to help determine if one or more of the authors’ claims are reliable, and 4) submitting a research compendium that documents the students' work in a reproducible format that includes the code and data used in the assignment.

We describe how to implement a replication report assignment suitable for upper-level undergraduates and graduate students in archaeology. Our experience is based on an upper-level archaeology class on stone artefact analysis taught during the Spring quarter of 2019 at the University of Washington. The class format includes a weekly cycle of lecture, discussion seminar, and hands-on laboratory activities. The assignments include seminar notes, lecture quizzes, laboratory worksheets, and two longer empirical reports. For the term that we report here, the class had 16 students and one graduate student teaching assistant. This is a typical size for this class, and similar to the usual size of upper-level laboratory classes in the archaeology program at the University of Washington. Our students are mostly in social science and humanities majors with varying levels of statistical competence. Here we survey the literature on similar types of assignments in other fields to identify common elements that other fields have identified as important principles and skills. We describe our assignment and discuss student feedback on our implementation. Finally, we offer recommendations for how to use replication reports to teach archaeology students.

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in @marwick2017open, we include all our assignment materials, as well as the entire R code used for all the analysis and visualizations contained in this paper, in our compendium at http://doi.org/10.17605/OSF.IO/DBSW9. Also in this version-controlled compendium are the raw data for the analyses reported here. The figures and results presented here can be independently reproduced with the code and data in this repository. In our compendium our code is released under the MIT license, our data as CC-0, and our figures, and assignment instructions and grading rubric as CC-BY, to enable maximum re-use [for more details, see @Marwick2017].

@@

@barba2018terminologies points out that although there has been prolific discussion of the terms "reproducibility" and "replication" in many disciplines in recent years, confusion and conflicting uses are widespread. In her survey of relevant literature, Barba finds that some fields make no distinction between "reproducibility" and "replication", and among fields that do recognise a distinction between the two, the meanings are sometimes directly inverted. Here we follow what Barba has identified as the most common, long-established and highly cited definitions of these terms, as also recently recommended by the National Academies of Sciences, Engineering, and Medicine [-@national2019reproducibility]. Reproducibility is the ability to obtain results by using the same data, code, and procedures provided by the original authors [@Marwick2017]. This is only possible when the authors make all those materials available, for example in a research compendium [@marwick2018packaging]. Replication is the ability to arrive at the same scientific conclusions in a new study, collecting new data (possibly with different methods) and completing new analyses.

In the following section we briefly survey replication assignments described in other fields to survey the variety of forms this usually takes. Replication assignments across different fields may not strictly fit into the above definition of replication because they do not always involve a completely new study. Nevertheless, we consider that if a study departs from any of the original materials, e.g. new data with previously published code, or new code with previously published data, then it is broadly within the definition of replication.

@@

This brief survey demonstrates that replication assignments are widely known in economics, psychology, political science, neuroscience and other fields [e.g. @roettger2018towards] . Common elements include group work, use of open source software and services to make the replication results openly accessible to anyone, and a scaffolded, stepwise approach to the task to ensure that students receive instructor support at multiple stages in the assignment. To the best of our knowledge replication assignments are not common in archaeology programs, although the tools and data structures are generally similar to other social sciences. We posted a message to the Society of American Archaeology Teaching Archaeology Interest Group e-community on 27 May 2019 to ask for examples of replication assignments used in teaching archaeology, and received no replies from anyone teaching with replication. More broadly in archaeology replication and reproducibility has received limited, but growing, attention. Elsewhere we have documented recent rapid increases in the number of publications that include code and data to enable readers to reproduce the published results [@Marwick_Schmidt_2019]

@@

Our main finding is that replication assignments are valued and in common use throughout the social sciences, and that they can be effective in teaching archaeology. Specifically, we found it possible to conduct a small scale replication assignment as part of an upper-level undergraduate archaeology class. Student feedback indicated that it was a valuable new experience for them, and for the discipline, even if they could not see themselves doing it again. We found that although this was a new and unconventional assignment, these elements did not have a negative effect on students' grades. Although our study is limited by its small size, when considered with the numerous other reports of replication assignments in other fields, we believe this approach will work in many types of archaeology classes.  Replication assignments have an important role in closing the gap between the ideals of archaeological science and preparing students to tackle the practical challenges of doing archaeological science transparently and reproducibly.

@@

```{r}
# what commit is this file at? 
git2r::repository(here::here())
```

>>>>

abstract: |
  The value of new archaeological knowledge is strongly determined by how credible it is, and a key measure of scientific credibility is how replicable new results are. However, few archaeologists learn the skills necessary to conduct replication as part of their training. This means there is a gap between the ideals of archaeological science and the skills we teach future researchers. Here, we argue for replications as a core type of class assignment in archaeology courses to close this gap and establish a culture of replication and reproducibility. We review replication assignments in other fields, and describe how to implement a replication assignment suitable for many types of archaeology programs. We describe our experience with replication in an upper-level undergraduate class on stone artefact analysis. Replication assignments can help archaeology programs give students skills that enable transparent and reproducible research.

@@

In his influential study of replication, sociologist of science Harry Collins argued that replication is at the core of scientific practice, writing “Replicability [...] is the Supreme Court of the scientific system” [@collins1992changing p. 19]. Like other observers of science, Collins claimed that in replication, private observations become communal facts, offering vital protection from error and fraud. In this paper we propose a new type of assignment for the archaeology classroom, the replication report, to better align the practice of teaching archaeology with the transparency and openness ideals of science [@nosek2015promoting]. The replication report assignment involves four steps: 1) students analysing a published report to determine the main claims made by the authors of that report, 2) students obtaining the data used by the authors, 3) analysing that data to help determine if one or more of the authors’ claims are reliable, and 4) submitting a research compendium that documents the students' work in a reproducible format including the code and data used in the assignment.

We describe how to implement a replication report assignment suitable for upper-level undergraduates and graduate students in archaeology. Our experience is based on an upper-level archaeology class on stone artefact analysis taught during the Spring quarter of 2019 at the University of Washington. The class format includes a weekly cycle of lecture, discussion seminar, and hands-on laboratory activities. The assignments include seminar notes, lecture quizzes, laboratory worksheets, and two longer empirical reports. For the term that we report here, the class had 16 students and one graduate student teaching assistant. This is a typical size for this class, and similar to the usual size of upper-level laboratory classes in the archaeology program at the University of Washington. Our students are mostly in social science and humanities majors with varying levels of statistical competence. Here, we survey the literature on similar types of assignments in other fields to identify common elements that other fields have identified as important principles and skills. We describe our assignment and discuss student feedback on our implementation. Finally, we offer recommendations for how to use replication reports to teach archaeology students.

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in @marwick2017open, we include all our assignment materials, as well as the entire R code used for all the analysis and visualizations contained in this paper, in our compendium at http://doi.org/10.17605/OSF.IO/DBSW9. Also in this version-controlled compendium are the raw data for the analyses reported here. The figures and results presented here can be independently reproduced with the code and data in this repository. In our compendium, our code is released under the MIT license, our data as CC-0, and our figures, and assignment instructions and grading rubric as CC-BY, to enable maximum re-use [for more details, see @Marwick2017].

@@

@barba2018terminologies points out that although there has been prolific discussion of the terms "reproducibility" and "replication" in many disciplines in recent years, confusion and conflicting uses are widespread. In her survey of relevant literature, Barba finds that some fields make no distinction between "reproducibility" and "replication." Among fields that do recognise a distinction between the two, the meanings are sometimes directly inverted. Here we follow what Barba has identified as the most common, long-established and highly cited definitions of these terms, as also recently recommended by the National Academies of Sciences, Engineering, and Medicine [-@national2019reproducibility]. Reproducibility is the ability to obtain results by using the same data, code, and procedures provided by the original authors [@Marwick2017]. This is only possible when the authors make all those materials available, for example in a research compendium [@marwick2018packaging]. Replication is the ability to arrive at the same scientific conclusions in a new study, collecting new data (possibly with different methods) and completing new analyses.

In the following section, we briefly survey replication assignments described in other fields to survey the variety of forms this usually takes. Replication assignments across different fields may not strictly fit into the above definition of replication because they do not always involve a completely new study. Nevertheless, we consider that if a study departs from any of the original materials, e.g. new data with previously published code, or new code with previously published data, then it is broadly within the definition of replication.

@@

This brief survey demonstrates that replication assignments are widely known in economics, psychology, political science, neuroscience and other fields [e.g. @roettger2018towards] . Common elements include group work, use of open source software and services to make the replication results openly accessible to anyone, and a scaffolded, stepwise approach to the task to ensure that students receive instructor support at multiple stages in the assignment. To the best of our knowledge, replication assignments are not common in archaeology programs, although the tools and data structures are generally similar to other social sciences. We posted a message to the Society of American Archaeology Teaching Archaeology Interest Group e-community on 27 May 2019 to ask for examples of replication assignments used in teaching archaeology, and received no replies from anyone teaching with replication. More broadly in archaeology, replication and reproducibility has received limited, but growing, attention. Elsewhere, we have documented recent rapid increases in the number of publications that include code and data to enable readers to reproduce the published results [@Marwick_Schmidt_2019]

@@

Our main finding is that replication assignments are valued and in common use throughout the social sciences, and that they can be effective in teaching archaeology. Specifically, we found it possible to conduct a small-scale replication assignment as part of an upper-level undergraduate archaeology class. Student feedback indicated that it was a valuable new experience for them, and for the discipline, even if they could not see themselves doing it again. We found that although this was a new and unconventional assignment, these elements did not have a negative effect on students' grades. Although our study is limited by its small size, when considered with the numerous other reports of replication assignments in other fields, we believe this approach will work in many types of archaeology classes.  Replication assignments have an important role in closing the gap between the ideals of archaeological science and preparing students to tackle the practical challenges of doing archaeological science transparently and reproducibly.

@@

```{r}
# what commit is this file at? 
if(git2r::in_repository(path = ".")) git2r::repository(here::here())
```
